---
title: "Fitness Data Analysis"
author: "BI"
date: "December 18, 2015"
output: html_document
---

The purpose of this analysis is to look at fitness data collected on personal fitness monitors in 5 minute increments. The data set analyzed is the Activity Monitoring Data in the csv file activity.csv.

We begin our analysis by loading the dataset and R packages we will be needing for the analysis.

```{r}
library(ggplot2)
library(dplyr)
library(xtable)
Data <- read.csv("activity.csv", header = TRUE)
```

Next I use the aggregate function to calculate the number of steps by date and then create a histogram of the result.

```{r}
spds <- aggregate(steps~date, data=Data, sum)
hist(spds$steps, ylab = "Number of Days", xlab = "Sum of Steps", main = "Steps per Day")
```

Next I look at the mean and median of the data on a daily basis and I get the following:

```{r}
tblave <- xtable(aggregate(steps~date, data=Data, mean))
tblmn <- xtable(aggregate(steps~date, data=Data, median))
print(tblave, type = "HTML")
print(tblmn, type = "HTML")
```
