---
title: "Fitness Data Analysis"
author: "BI"
date: "December 18, 2015"
output: html_document
---

The purpose of this analysis is to look at fitness data collected on personal fitness monitors in 5 minute increments. The data set analyzed is the Activity Monitoring Data in the csv file activity.csv.

We begin our analysis by loading the dataset and R packages we will be needing for the analysis.

```{r}
library(ggplot2)
library(dplyr)
library(xtable)
Data <- read.csv("activity.csv", header = TRUE)
```

Next I use the aggregate function to calculate the number of steps by date and then create a histogram of the result.

```{r}
spds <- aggregate(steps~date, data=Data, sum)
hist(spds$steps, ylab = "Number of Days", xlab = "Sum of Steps", main = "Steps per Day")
```

Next I look at the mean data on a daily basis and I get the following:

```{r}
aggregate(steps~date, data=Data, mean)
```

When we print out the median we the result is so small that we get a lot of 0s as a result. Here are a few rows to demonstrate.

```{r}
demo <- aggregate(steps~date, data=Data, median)
demo[1:5, ]
```

Looking at the data set it is easy to see why. The data is overwhelmed by the number of zeros.Of 17,568 observations 11,014 are 0

```{r}
sum(Data$steps == 0, na.rm = TRUE)
```

I chose to groom the data to give a more meaningful result. I subsetted for the data for values that were not NA or 0 and calculated the median on the result. This is not a daily median, but a media for the entire data set less zeros and NAs. I then calculate the median of the average daily number of steps.

```{r}
mdsna <- Data[,1]
mdsna <- mdsna[!is.na(mdsna)]
mdsna <- mdsna[mdsna != 0]
md <- median(mdsna)
md
mdd <- aggregate(steps~date, data=Data, mean)
median(mdd$steps)
```

I use the aggregate function to create a new data set of mean steps and by interval and plot the result and return the maximum value by interval
```{r}
spd <- aggregate(steps~interval, data=Data, mean)
plot(spd$interval, spd$steps, type = "l")
spdmx <- spd[order(-spd$steps),]
spdmx[1,]
```
